\Large \bf Cyclone: High Availability for Persistent Key Value Stores

### 2017-08-03 paper/paper.tex
Updates.

DPDK does not by itself provide a TCP stack. This is not a problem. RAFT (and
indeed most consensus protocols) tolerate network losses and reordering by
design due to the need to support asynchronous communication. In addition, most
datacenter networks rarely drop or reorder packets and provide full bisection
bandwith between servers that might serve as Cyclone replicas. We therefore
jettisoned TCP and chose to send raw IP packets encapsulated in ethernet
frames. We currently follow this communication model both for server to server
communication between replicas as well as client server communication for
requests and responses. Although a detailed evaluation is made later, switching
from the kernel TCP/IP stack to DPDK reduces the latency between machines in our
testbed from {\tt 18 us} to as low as {\tt 5 us}, providing a significant boost
to performance.

@@

This entire process is done without making any copies of the received packet.
To illustrate how this is achieved, Figure~\ref{fig:packet_layout} shows how
Cyclone manipulates packet layouts across the two steps of prepending a RAFT
header and transmitting to follower replicas. DPDK describes packets using an
``mbuf'' data structure. Roughly speaking, an ``mbuf'' consists of a flat array
of bytes actually containing the packet and a fixed size piece of external
metadata that describes various aspects of the packet, most crucially a pointer
to the start and end of the packet in the byte array. DPDK's userspace drivers
receive packets from the NIC such that they are offset in the byte array by a
configurable amount referred to as ``headroom''. We strip off the existing
network headers in the packet and prepend RAFT related information specific to
each log entry in the headroom by shifting the start pointer appropriately.
These operations are standard enough for software packet switches that DPDK
provides convenient library calls for it. For the final step, we need to prepare
the packet for transmission to the various follower replicas. To do this we
prepare a different packet containing an ethernet header for \emph{each}
targeted replica and ``chain'' the data packet to each of these headers. Each
header is then separately handed off to the driver for transmission via the NIC,
carrying the data packet with it by association.

@@

\subsection{Ganged Operations}
Supporting reads or writes to multiple keys that map to different physical logs
requires a ganged operation across those logs such that all the operations
complete together in the face of failures. Since our starting point is a single
key value store, we constrained Cyclone's design early on to avoid the need for
any two phase commit.

The first key problem we need to deal with is that the leaders for the different
RAFT instances can be located on different machines. To avoid having a client
co-ordinate different machines for a single operation, we constrain the RAFT
leader election algorithm for the different RAFT instances to converge to the
same machine. We do this by triggering re-elections for a RAFT instance other
than that of the first physical log until it is on the same machine as the
\emph{first} physical log. This process is therefore resilient to failed
machines. RAFT requires that the leader have the most up to date log. To ensure
that the process converges we therefore ensure that the currently elected leader
brings a majority quroum up to date before triggering a relection. We also do
not accept any requests from clients until all leaders are co-located. This
process is guaranteed to converge as long as all no network partitions occur
\emph{within a machine} i.e. some RAFT instances are able to send and receive
packets, while others are not, a possiblity we discount due to our single
process design.

@@

\begin{figure*}
  \centering
\begin{verbatim}
    // Apply ganged operation
    event_apply_ganged_op (packet, barrier)
    {
      if(co-ordinator)
        atomic set bit me in barrier.mask         
        do
         for each participant p in operation
          if public_data[p].view > packet.view
           barrier.failed = true 
           atomic set bit for p in barrier.mask
        while barrrer.mask != mask of all participants
        if barrier.failed
         Ganged replication failed. Send retry to client. return.
        else       
         execute operation
         send response
      else
        wait until 
         public_data[co-ordinator].view > packet.view OR
         barrier.mask == mask of participating KV cores
        if public_data[co-ordinator].view > packet.view OR
           barrier.failed
         Ganged replication failed. Send retry to client. return.
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure*}

>>>>

DPDK does not by itself provide a TCP stack. This is not a problem. RAFT (and
indeed most consensus protocols) tolerate network losses and reordering by
design due to the need to support asynchronous communication. In addition, most
datacenter networks rarely drop or reorder packets and provide full bisection
bandwith between servers that might serve as Cyclone replicas. We therefore
jettisoned TCP and chose to send raw IP packets encapsulated in ethernet
frames. Cyclone takes complete control of an ethernet interface, receiving all
packets directed to it, while the IP and ethernet addresses provide sufficient
information to route the packet if necessary through multiple switches. We
currently follow this communication model both for server to server
communication between replicas as well as client server communication for
requests and responses. Although a detailed evaluation is made later, switching
from the kernel TCP/IP stack to DPDK reduces the latency between machines in our
testbed from {\tt 18 us} to as low as {\tt 5 us}, providing a significant boost
to performance.

@@

This entire process is done without making any copies of the received packet.
To illustrate how this is achieved, Figure~\ref{fig:packet_layout} shows how
Cyclone manipulates packet layouts across the two steps of prepending a RAFT
header and transmitting to follower replicas. DPDK describes packets using an
``mbuf'' data structure. Roughly speaking, an ``mbuf'' consists of a flat array
of bytes actually containing the packet and a fixed size piece of external
metadata that describes various aspects of the packet, most crucially a pointer
to the start and end of the packet in the byte array. DPDK's userspace drivers
receive packets from the NIC such that they are offset in the byte array by a
configurable amount referred to as ``headroom''. We strip off the existing
network headers in the packet and prepend RAFT related information specific to
each log entry in the headroom by shifting the start pointer appropriately.
These operations are standard enough for software packet switches that DPDK
provides convenient library calls for it. For the final step, we need to prepare
the packet for transmission to the various follower replicas. To do this we
prepare a different packet containing a network header for \emph{each}
targeted replica and ``chain'' the data packet to each of these headers. Each
header is then separately handed off to the driver for transmission via the NIC,
carrying the data packet with it by association.

@@

\subsection{Ganged Operations}
Supporting reads or writes to multiple keys that map to different physical logs
requires a ganged operation across those logs such that all the operations
complete together in the face of failures. One option is to have an external
co-ordinator run a distributed transaction using two phase commit. In addition
to requiring extra external resources to host the co-ordinator one must also
worry about failure recovery. In addition this requires two round trips to the
leader to complete what should have been a single batched write request from a
client that is stateless and allowed to fail - such as with key value stores
like Rocksdb~\cite{rocksdb}. We therefore describe a simpler and more performant
solution called ganged operations integrated into Cyclone. This allows a client
to issue a batch of {\tt PUT} operations as a gang simultaneously and atomically
updating a batch of keys in a single call.

The first key problem we need to deal with is that the leaders for the different
RAFT instances to which the keys map can be located on different machines. To
avoid having a client co-ordinate different machines for a single operation, we
constrain the RAFT leader election algorithm for the different RAFT instances to
converge to the same machine. We do this by triggering re-elections for a RAFT
instance other than that of the first physical log until it is on the same
machine as the \emph{first} physical log. This process is therefore resilient to
failed machines. RAFT requires that the leader have the most up to date log. To
ensure that the process converges we therefore ensure that the currently elected
leader brings a majority quroum up to date before triggering a relection. We
also do not accept any requests from clients until all leaders are
co-located. We assume that no network partitions occur \emph{within a machine}
i.e. some RAFT instances are able to send and receive packets, while others are
not, a possiblity we discount due to our single process design.

@@

\begin{figure*}
  \centering
\begin{verbatim}
    // Apply ganged operation
    event_apply_ganged_op (packet, barrier)
    {
      if(co-ordinator)
        atomic set bit me in barrier.mask         
        do
         for each participant p in operation
          if public_data[p].view > packet.view
           barrier.failed = true 
           atomic set bit for p in barrier.mask
        while barrrer.mask != mask of all participants
        if barrier.failed
         Ganged replication failed. Send retry to client. return.
        else       
         execute operation
         send response
      else
        wait until 
         public_data[co-ordinator].view > packet.view OR
         barrier.mask == mask of participating cores
        if public_data[co-ordinator].view > packet.view OR
           barrier.failed
         Ganged replication failed. Send retry to client. return.
        else
         execute operation
    }
\end{verbatim}
\caption{Ganged Operation}
\label{fig:ganged_ops}
\end{figure*}
